---
output:
  word_document: default
  html_document: default
---
# MIS503 – Final Project
## Badillo, Roy
### Zillow Home Value Index Analysis

# Wake County Home Sales

```{r}
library(tidyverse)
library(readr)
SingleFamilyResidenceRental <- read_csv("SingleFamilyResidenceRental.csv")
View(SingleFamilyResidenceRental)

SingleFamilyResidenceSales <- read_csv("SingleFamilyResidenceSales.csv")
View(SingleFamilyResidenceSales)




```


```{r}

WakeCountySales <- select(SingleFamilyResidenceSales, CountyName, State)
WakeCountySales <- filter(SingleFamilyResidenceSales, State == "NC")
WakeCountySales <- filter(SingleFamilyResidenceSales, CountyName == "Wake County")

WakeCountySales <- select(WakeCountySales, c(RegionName, State, CountyName, Metro, "1996-05", "1997-05", "1998-05", "1999-05", "2000-05", "2001-05", "2002-05", "2003-05", "2004-05","2005-05", "2006-05","2007-05", "2008-05", "2009-05", "2010-05", "2011-05", "2012-05", "2013-05", "2014-05", "2015-05", "2016-05", "2017-05", "2018-05"))

WakeCountySales <- WakeCountySales %>%
  rename( "1996" = "1996-05", "1997" = "1997-05" , "1998" = "1998-05", "1999" = "1999-05", "2000" = "2000-05" , "2001" = "2001-05", "2002" = "2002-05", "2003" = "2003-05", "2004" = "2004-05", "2005" = "2005-05", "2006" = "2006-05", "2007" = "2007-05", "2008" = "2008-05", "2009" = "2009-05", "2010" = "2010-05", "2011" = "2011-05", "2012" = "2012-05", "2013" = "2013-05", "2014" = "2014-05", "2015" = "2015-05", "2016" = "2016-05", "2017" = "2017-05", "2018" = "2018-05")

WakeCountySales <- WakeCountySales %>%
  pivot_longer(cols = c("1996", "1997", "1998", "1999", "2000", "2001", "2002", "2003", "2004","2005", "2006", "2007", "2008", "2009", "2010", "2011", "2012", "2013", "2014", "2015", "2016", "2017", "2018"), names_to = 'YR', values_to = 'ZHVI')

ggplot(WakeCountySales, aes(YR, ZHVI, colour = RegionName)) + geom_point() + theme(axis.text.x = element_text(angle = 90, vjust=0.5)) + scale_y_continuous(name="Home Values (in $)", labels = scales::comma) + ggtitle("Wake County Home Sales")



```


```{r}
ggplot(WakeCountySales, aes(YR, ZHVI, colour = RegionName)) + geom_col() + theme(axis.text.x = element_text(angle = 90, vjust=0.5)) + scale_y_continuous(name="Home Values (in $)", labels = scales::comma) + ggtitle("Wake County Home Sales")

```


a. What have been the overall trends in Wake County Home Values?

- They have all appreciated.

b. There were dips in home values in the past 20 years. What years did these occur?
- The dips started in 1997 and 2008 but the bottom of the dips were in 1998 and 2012.


c. Based on the analysis, where would be the least expensive area to purchase home? Most expensive area?
- Zebulon is the least expensive and New Hill the most expensive (huge increase from 2017-2018)

d. Are any area home values trending down? Is there one area that stands out compared to others?

- No areas are trending down but some areas are certainly increasing in price faster than others for example Rolesville has incresed 32% in value while New Hill has appreciated 117% in the same period between 1996 and 2018.



# NC Rental Market


```{r}
SingleFamilyResidenceRental <- read_csv("SingleFamilyResidenceRental.csv")
View(SingleFamilyResidenceRental)
Rentals <- filter(SingleFamilyResidenceRental, State == "NC")
Rentals <- filter(SingleFamilyResidenceRental, RegionName %in%
                    c("Asheville", "Charlotte", "Durham", "Fayetteville", "Raleigh", "Wilmington"))

Rentals <- select(Rentals, c(RegionName, State, "2010-11", "2011-11", "2012-11", "2013-11", "2014-11", "2015-11", "2016-11", "2017-11", "2018-10"))

Rentals <- Rentals %>%
  rename( "2010" = "2010-11", "2011" = "2011-11", "2012" = "2012-11", "2013" = "2013-11", "2014" = "2014-11", "2015" ="2015-11", "2016" = "2016-11",  "2017" = "2017-11",  "2018" = "2018-10")


Rentals <- Rentals %>%
  pivot_longer(cols = c("2010", "2011", "2012", "2013", "2014", "2015", "2016",  "2017",  "2018"), names_to = 'YR', values_to = 'ZHVI')

Rentals <- filter(Rentals, State == "NC")

ggplot(Rentals, aes(YR, ZHVI, colour = RegionName)) + geom_point() + theme(axis.text.x = element_text(angle = 90, vjust=0.5)) + scale_y_continuous(name="Average Rental Cost", labels = scales::comma) + ggtitle("Rental Costs in NC Cities")

```
```{r}
ggplot(Rentals, aes(YR, ZHVI, colour = RegionName)) + geom_col() + theme(axis.text.x = element_text(angle = 90, vjust=0.5)) + scale_y_continuous(name="Average Rental Cost", labels = scales::comma) + ggtitle("Rental Costs in NC Cities")
```
a. What has been the overall trend in the rental market around the state? Are there any cities that have not followed this trend?
-The overall trend is an increase in rent however, Fayetteville has actually decreased since 2010.

b. Where is the most expensive city to rent in? Least expensive? 
 Most expensive was Raleigh for a while but Asheville took the lead in 2017 for most expensive, least expensive is Fayetteville


c. You are trying decide between Wilmington and Asheville. Which market has the lowest rent?

- Wilmington has lower rent than Asheville (and beaches)


# Home Values in Select Rental Markets

```{r}
NCHomeSales <- select(SingleFamilyResidenceSales, c(RegionName, State, Metro, "1996-05", "1997-05", "1998-05", "1999-05", "2000-05", "2001-05", "2002-05", "2003-05", "2004-05","2005-05", "2006-05","2007-05", "2008-05", "2009-05", "2010-05", "2011-05", "2012-05", "2013-05", "2014-05", "2015-05", "2016-05", "2017-05", "2018-05"))

NCHomeSales <- NCHomeSales %>%
  rename( "1996" = "1996-05", "1997" = "1997-05" , "1998" = "1998-05", "1999" = "1999-05", "2000" = "2000-05" , "2001" = "2001-05", "2002" = "2002-05", "2003" = "2003-05", "2004" = "2004-05", "2005" = "2005-05", "2006" = "2006-05", "2007" = "2007-05", "2008" = "2008-05", "2009" = "2009-05", "2010" = "2010-05", "2011" = "2011-05", "2012" = "2012-05", "2013" = "2013-05", "2014" = "2014-05", "2015" = "2015-05", "2016" = "2016-05", "2017" = "2017-05", "2018" = "2018-05")

NCHomeSales<- filter(NCHomeSales, State == "NC")

NCHomeSales <- filter(NCHomeSales, Metro %in%
                    c("Asheville", "Charlotte-Concord-Gastonia",
"Raleigh", "Wilmington"))

NCHomeSales <- NCHomeSales %>%
  pivot_longer(cols = c("1996", "1997", "1998", "1999", "2000", "2001", "2002", "2003", "2004","2005", "2006", "2007", "2008", "2009", "2010", "2011", "2012", "2013", "2014", "2015", "2016", "2017", "2018"), names_to = 'YR', values_to = 'ZHVI')

ggplot(NCHomeSales, aes(Metro, ZHVI)) + geom_violin() + geom_boxplot(width=0.1)


```



a. According to the results, which market has the lowest median price (represented as horizontal bar in box plot)?
- Charlotte-Concord-Gastonia

b. The violin plot will show density meaning the wider the plot is, the more observations occur within that area. Which market has the most density around the median value of homes?

- Charlotte-Concord-Gastonia

# Relocation Home Value Comparison


```{r}
NationalHomeSales <- select(SingleFamilyResidenceSales, c(RegionName, State, Metro, "1996-05", "1997-05", "1998-05", "1999-05", "2000-05", "2001-05", "2002-05", "2003-05", "2004-05","2005-05", "2006-05","2007-05", "2008-05", "2009-05", "2010-05", "2011-05", "2012-05", "2013-05", "2014-05", "2015-05", "2016-05", "2017-05", "2018-05"))
```


```{r}
NationalHomeSales<- NationalHomeSales %>%
  rename( "1996" = "1996-05", "1997" = "1997-05" , "1998" = "1998-05", "1999" = "1999-05", "2000" = "2000-05" , "2001" = "2001-05", "2002" = "2002-05", "2003" = "2003-05", "2004" = "2004-05", "2005" = "2005-05", "2006" = "2006-05", "2007" = "2007-05", "2008" = "2008-05", "2009" = "2009-05", "2010" = "2010-05", "2011" = "2011-05", "2012" = "2012-05", "2013" = "2013-05", "2014" = "2014-05", "2015" = "2015-05", "2016" = "2016-05", "2017" = "2017-05", "2018" = "2018-05")
```


```{r}
NationalHomeSales <- NationalHomeSales %>%
  pivot_longer(cols = c("1996", "1997", "1998", "1999", "2000", "2001", "2002", "2003", "2004","2005", "2006", "2007", "2008", "2009", "2010", "2011", "2012", "2013", "2014", "2015", "2016", "2017", "2018"), names_to = 'YR', values_to = 'ZHVI')

NationalHomeSales <- filter(NationalHomeSales, Metro %in%
                    c("New York-Newark-Jersey City", "Houston-The Woodlands-Sugar Land",
"Denver-Aurora-Lakewood", "Chicago-Naperville-Elgin"))
```


```{r}
ggplot(NationalHomeSales, aes(YR, ZHVI)) + geom_point() + facet_wrap(~Metro) + theme(axis.text.x = element_text(angle = 90, vjust=0.5, size=5)) + scale_y_continuous(name="Home Values (in $)", labels = scales::comma) +  ggtitle("City Home Value Comparison")
```

1. Based on your analysis, which city’s housing is most affordable? Least affordable?

-Houston is most affordable, New York is least affordable.

2. Which cities saw the largest change in prices over the past 5 years? Which city has remained more consistent (i.e., no huge swings up or down in home values)?
- New York and Colorado have seen the biggest swing in value, Houston remained the most consistent.


3. During the market downturn in 2012, which cities were most impacted? Which cities have recovered?

- New York City was the most affected during 2012 followed by Chicago, New York has recovered but Chicago has not.





```{
```


